RCloud: Integrating Exploratory Visualization, Analysis and Deployment

### 2014-03-31 system.tex
stab at the web architecture / interactive / deployment system stuff

HTTP and web services are convenient but much of the protocol is
stateless. (For example, GET requests are required to not change the
remote state, and results can be cached in transparent proxies along
the network). In the case of close communication between a running
R process and a web browser, we need something that is more flexible
than distinct URLs for every R process (with the resulting
complication in providing access

%% \subsection{Technologies: R, Python, HTML5, interactive notebooks, etc.}

@@

%% A key engineering decision in RCloud was to rely on full two-way
%% communication between the web client human interface, and data and
%% computing resources in the cloud. It is easier to engineer a design
%% in which the human interface is pushed one-way from the server
%% and thereafter all interaction takes place within the client.
%% Two-way communication is necessary, though, where the size of
%% the data or computational performance makes it impractical to
%% move all computation to the client.

@@

Notebooks as versioned subroutines, web services.

>>>>

HTTP and web services are convenient but much of the protocol is
stateless. (For example, GET requests are required to not change the
remote state, and results can be cached in transparent proxies along
the network). In the case of close communication between a running
R process and a web browser, we need something that is more flexible
than distinct URLs for every R process, and which embraces
statefulness of both calculations and insights in progress.

\subsection{Interactive notebooks, etc.}

@@

A key engineering decision in RCloud was to rely on full two-way
communication between the web client human interface, and data and
computing resources in the cloud. It is easier to engineer a design
in which the human interface is pushed one-way from the server
and thereafter all interaction takes place within the client.
Two-way communication is necessary, though, where the size of
the data or computational performance makes it impractical to
move all computation to the client.

In addition, R is naturally better as a language for analysis, and
JavaScript as a language for interactive visualization.  To draw on
the strengths of both languages and both environments, the connection
between the languages is not just procedural: it makes \emph{closures}
and \emph{first-class functions} available across the network link.
This provides great flexibility, so that for example, a chart built
with dc.js or leaflet.js can call back to analysis functions in R
without having to formalize the protocol between the processes.

@@

Because notebooks are continually published, any work is immediately
available both as a subroutine and visual component. Close colleagues
can start on the next stage of analysis, or delve into the data,
even while the original author is polishing the algorithm or its
presentation. The code is the page, and can either be shaped into
a function of inputs and outputs, or the linear cells of the notebook
can be reworked into a full-fledged HTML layout.
